\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-03-29 paper.tex

Historically, \(k-D\) representations of \(p-D\) data have been computed
using multidimensional scaling (MDS), which includes principal
components analysis (PCA) as a special case. The \(k-D\) representation
can be considered to be a layout of points in \(k-D\) produced by an
embedding procedure that maps the data from \(p-D\). In MDS, the \(k-D\)
layout is constructed by minimizing a stress function that differences
distances between points in \(p-D\) with potential distances between
points in \(k-D\). Various formulations of the stress function result in
non-metric scaling \citep{article62} and isomap \citep{article63}.
Challenges in working with high-dimensional data, including
visualization, are outlined in \citet{Johnstone2009}.

>>>>

Historically, \(k-D\) representations of \(p-D\) data have been computed
using multidimensional Scaling (MDS) \citep{borg2005}, which includes
principal components analysis (PCA) \citep{jolliffe2011} as a special
case. The \(k-D\) representation can be considered to be a layout of
points in \(k-D\) produced by an embedding procedure that maps the data
from \(p-D\). In MDS, the \(k-D\) layout is constructed by minimizing a
stress function that differences distances between points in \(p-D\)
with potential distances between points in \(k-D\). Various formulations
of the stress function result in non-metric scaling \citep{kruskal1964}
and isomap \citep{silva2002}. Challenges in working with
high-dimensional data, including visualization, are outlined in
\citet{johnstone2009}.
